---
title: "Project: Data Cleaning and QC"
author: "Jan Knappe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
toc_depth: 3
number_sections: true
theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(mice)
library(mvoutlier, quietly = TRUE)
source("dd.plot2.R")
```


## Data import

```{r}
data = 
    read_csv(
        "data/weather_data.csv",
        col_types = cols(date = col_date(format = ""),
                         airTemperature = col_double(),
                         atmPressure = col_double(),
                         netRadiation = col_double(),
                         rainfall = col_double(),
                         relativeHumidity = col_double(),
                         soilTemperature = col_double())
    )
```

## Summary stats

```{r}
data %>% summary()
```
```{r}
data %>% glimpse()

```

## Data manipulation

Strange entries in soilTemperature 

```{r}
summary(data$soilTemperature)
ggplot(data, aes(x = date, y = soilTemperature)) + 
    geom_point()

length(data$soilTemperature[data$soilTemperature < - 50])

# replace everything below -50 degree with NA
(data <- data %>% mutate_at("soilTemperature", ~ifelse(. < -50 , NA, .)))
```


## Pattern of missing data

```{r}
misspat <- mice::md.pattern(data)

mtext("N. missing values in variable", side = 1, padj = 5)
  mtext("N. observations with pattern", side = 2, padj = -2)
  mtext("Variable", side = 3, padj = -3)
  mtext("N. missing values in pattern", side = 4, padj = -2)

```


Column relativeHumidity contains NA's

```{r}
ggplot(data, aes(x = date, y = relativeHumidity)) + 
    geom_point()

```

## Linear interpolation

```{r}
ina <- which(is.na(data$relativeHumidity))
dim(data[ina,])
#datafilt <- subset(data, filtration =="yes")

ppm3int <- approx(x = data$date, # helper variable: time
                  y = data$relativeHumidity, # variable with missing values
                  xout = data$date[ina]) # which times to interpolate
ppm3int

```

```{r}
plot(relativeHumidity ~ date, data = data, pch = 16,
     xlab = "Time (days)", ylab = "Relative Humidity")

# Add the interpolated values
points(ppm3int$x, ppm3int$y, pch = 16, col = "red")

```

## Outliers with Mahalanobis distance

```{r}
dat.noNA <- na.exclude(data)

# Calculate Mahalanobis distance
centre <- colMeans(dat.noNA[,c("airTemperature","atmPressure","netRadiation","rainfall", "relativeHumidity", "soilTemperature")])
varcov <- cov(dat.noNA[,c("airTemperature","atmPressure","netRadiation","rainfall", "relativeHumidity", "soilTemperature")])
# mahalanobis() gives the squared distance by default so we take the square-root
mdist <- sqrt(mahalanobis(dat.noNA[,c("airTemperature","atmPressure","netRadiation","rainfall", "relativeHumidity", "soilTemperature")], 
            center = centre, cov = varcov))

# Visualise the distances
hist(mdist, col = "grey", xlab = "Mahalanobis distances", main = "")

```
```{r}
dout <- qchisq(p = 0.01, lower.tail = FALSE,
               df =6) # df: number of variables used to calculate the distance
dout

any(mdist > dout)
```


## Outliers with Robust Mahalanobis distance

```{r}
# Calculate the robust Mahalanobis distance
# and identify outliers
d <- dd.plot2(dat.noNA[,c("airTemperature","atmPressure","netRadiation","rainfall", "relativeHumidity", "soilTemperature")],
             alpha = 0.01)

# Inspect the robust Mahalanobis distances
hist(d$md.rob, col = "grey", xlab = "Robust Mahalanobis distances", main = "")

index <- which(d$outliers)
dat.noNA[index,]
```

```{r}
cols <- rep(grey(0,0.5), nrow(dat.noNA))
cols[index] <- "red"
pairs(dat.noNA[,c("airTemperature","atmPressure","netRadiation","rainfall", "relativeHumidity", "soilTemperature")],
      pch = 16, col = cols)

```


so these are all high rain fall events!
some might be indeed a little extreme, but everything above ~4 mm?


```{r}

dat.noNA$rMhlnb_out <- d$outliers

dat.noNA %>% #filter(date > '2015-12-31'  & date < '2018-01-01') %>%
ggplot(aes(x = date, y = rainfall, color = rMhlnb_out)) + 
  geom_point()
```




```{r}
dat.noNA %>% pivot_longer(cols = -c(date, rMhlnb_out, numdate),
                          names_to = "variable") %>% 
ggplot(aes(x = date, y = value, )) + 
  facet_grid(variable~., scales = "free_y") +
  geom_point()
```


Alle anderen Variablen haben eine mehr oder weniger offensichtliche Saisonalität und dadurch eine deutlichere Normalverteilung:

```{r}
dat.noNA %>% pivot_longer(cols = -c(date, rMhlnb_out, numdate),
                          names_to = "variable") %>% 
ggplot(aes(x = value, )) + 
  facet_grid(.~variable, scales = "free_x") +
  geom_histogram()
```

Gut, ne. Die Bodentemperatur ist ja fast uniform, Strahlung und Luftfeuchte deutlich skewed...



```{r}
# Load the package
library(EMD)

# Create a numeric time variable from the dates
dat.noNA$numdate <- as.numeric(difftime(dat.noNA$date, min(dat.noNA$date), units = "days"))

# Run the EMD and extract the Intrinsic Mode Functions
# (we set the maximum number of IMF's to 10)
imfs <- emd(xt = dat.noNA$atmPressure, tt = dat.noNA$numdate, max.imf = 10)
imfs <- imfs$imf

#Let’s visualise the fitted IMF’s.
```


```{r}
# Visualise the IMF's
layout(matrix(1:ncol(imfs), ncol = 1))
par(mar = c(2,4,2,2))
apply(imfs, 2, plot, type = "l", col = "blue", xlab = "", ylab = "")
```
```{r}
# Make a histogram of the noise
hist(imfs[,1], col = "blue")
```


```{r}

```

